---
title: "Sylamer tools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sylamer tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Instalation and Setup

```{r setup}
#install library
#devtools::install_github("ipatop/SylamerTools")
library(SylamerTools)
```

# Description

This package contains three functions that can be used to aid with running [Sylamer](https://github.com/micans/sylamer) Stijn van Dongen, Cei Abreu-Goodger & Anton J. Enright; Nature Methods(2008), PMID: 18978784.

It contains:

1. A function to sort genes by any numerical condition
2. A function to create a shell commando to run Sylamer
3. A function to plot the Sylamer output

Sylamer is a tool that analyzes expression data to identify microRNA targets and small interfering (si)RNA off-target signals. It takes as input a ranked list of 3' untranslated regions (UTRs) that have been upregulated or downregulated following an miRNA or RNAi experiment, and generates a landscape plot showing the occurrence biases of various words across the gene ranking using hypergeometric P-values. 

# Sort genes 

Sylamer requires as input a sorted list of genes from up to down regulated. The function `sort_forSylamer` will sort any list of genes by any column name. 

Input requirements: a column named "gene", another column to sort genes by. Any other column will be ignored.

<table>

gene      |  log2FoldChange |
--------  | --------------- | 
mthl8     | 0.05805309      |	
--------  | --------------- | 
Lsp1gamma | 2.09023047      |	

</table>

Output: a file with the sorted genes ready to input into sylamer

**Input example**
```{r}
#This is how the input looks like
head(read.table(file = paste0(system.file("extdata", package = "SylamerTools"),"/res.txt"),header = T))
```
**Output example**

```{r}
sort_forSylamer(DEres = paste0(system.file("extdata", package = "SylamerTools"),"/res.txt"),sortBy = "log2FoldChange",outName =paste0(system.file("extdata", package = "SylamerTools") ,"/ForSylamer.txt"))

head(read.table(file = paste0(system.file("extdata", package = "SylamerTools"),"/ForSylamer.txt") ,header = T))
```

# Create sylamer command

The function `create_sylamer_script` will return a file ready to run.

The output from this function can be run in unix as follows:

```sh
sh sylamer_command.sh
```

This function takes the following options.

+ `utrFile`: File with the UTR sequences. Defauls is "dm3_flybase_3utrs.fa"
+ `kSylamer`: Oligo letter length or Word size, has to be the same length as those in the "words" file. Default is 6
+ `kMarkov`: Size of a smaller word to be used for correcting composition biases. Default is 4
+ `winSize`: How many sequences should be added in each consecutive window. Default is 200
+ `sylamer`: Sylamer executable path. Default is "/opt/linux64bin/sylamer"
+ `sylOutput`: Name of output file to write table in. Default is "shRNA_sylamer.output.tab"
+ `imageOut`: Name of output file to write plot in. Default is "shRNA_image.output.pdf"
+ `DEsorted`: Name of one or multiple list of sorted genes to input in sylamen. Default is c("ForSylamer.txt","ForSylamer_2.txt")
+ `extras`: Any extra parametere to pass into Sylamer. Default is "-a 10 -aa 5 --funny-ok"
+ `overwrite`: Boolean. If True, the script will overwrite any existing script with the same name. Default is T
+ `sylamer_comand`: Name of the file to write the command into "sylamer_command.sh"

```{r}
create_sylamer_script(utrFile =  paste0(system.file("extdata", package = "SylamerTools"),"/dm3_flybase_3utrs.fa"), overwrite = T, DEsorted=paste0(system.file("extdata", package = "SylamerTools"),c("/ForSylamer.txt","/ForSylamer_2.txt")))
```

**Output**

```{r}
read.delim("sylamer_command.sh",header = F)
```

# Plot Sylamer output

Sylamer output has to be organized to be able to interpret and plot them. To this end, we wrote a function thatsort the results and plot them. 

It takes the following options:

+ `oligosChosen`: Sequences to highlight (Any particular words that you want highlighted) Defaults is =c("TGTAAA","GATGCT")
+ `sylamer.out.tab`: Name of sylamer output to read from. Dafault is "shRNA_sylamer.output.tab"
+ `topOligos`: Number of top oligos to plot. How many best words should I highlight in the image? Defauls is 0
+ `lowOligos`: Number of low oligos to plot. Defauls is 0
+ `outName`: name for the plot output. Default is "sylamerplot"
+ `greylines`: Boolean, To plot or not greay background oligos. Default is F
+ `kSylamer`: Default is 6

**Sylamer Output**

```{r}
read.delim(paste0(system.file("extdata", package = "SylamerTools"),"/L2FcSh_sylamer.output.tab"))[1:3,1:3]
```
**Plot**

```{r}
sylamerplots(oligosChosen=c("TGTAAA","GATGCT"),sylamer.out.tab=paste0(system.file("extdata", package = "SylamerTools"),"/L2FcSh_sylamer.output.tab"),outName ="shRNA_fake")
```

```{r}
sylamerplots(oligosChosen=c("TGTAAA","GATGCT"),sylamer.out.tab=paste0(system.file("extdata", package = "SylamerTools"),"/L2FcSh_sylamer.output.tab"),greylines = T,outName = "shRNA_fake_markingGrey",free_ylim = T)
```
